---
title: "Composite Risk Difference with VA-CT-TMLE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Composite Risk Difference with VA-CT-TMLE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup}
library(vacttmle)
```

This package estimates a visit-aligned continuous-time TMLE for the composite
survival target

```text
P(T^d > tau, D^d > tau).
```

The reported risk difference is

```text
{1 - P(T^d1 > tau, D^d1 > tau)} -
{1 - P(T^d0 > tau, D^d0 > tau)}.
```

The intercurrent event `D` is part of the composite endpoint in this package
version. It is not a censoring process.

```{r}
dat <- simulate_va_trial(n = 500, scenario = "S1", seed = 1)
fit <- va_ct_tmle(dat, M = 20, B = 10, max_iter = 5, seed = 2)
fit
```

Diagnostics include component-wise EIF means and cumulative-weight summaries.

```{r}
fit$diagnostics
```

For trial data, use `as_va_data()` to adapt subject-level and person-interval
tables, then call `validate_va_data()` before estimation.
