---
title: "ITT and per-protocol estimands under treatment switching"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ITT and per-protocol estimands under treatment switching}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(vacttmle)
```

## The problem

In many randomized trials the treatment assigned at baseline is not the
treatment received throughout follow-up: control-arm patients cross over to the
experimental treatment, patients switch off toxic regimens, and others are lost
to follow-up. These events happen in continuous time and are recorded to the
day, while the covariates that drive them (disease status, biomarkers) are
measured only at scheduled visits.

Two estimands answer different questions, both mapped to ICH E9(R1) strategies:

- **Treatment-policy (ITT):** contrast the randomized arms leaving switching as
  it naturally occurs (censoring removed).
- **Hypothetical no-switching (per-protocol, PP):** contrast "assigned to arm
  *a* and never switch" (switching and censoring removed).

`vacttmle` estimates both from the exact switch and dropout dates, adjusting for
the visit-measured time-varying confounders that drive switching and dropout.

## A simulated trial

```{r}
dat <- simulate_va_switch_trial(n = 1000, scenario = "A2", seed = 1)
```

Scenario `"A2"` has crossover driven by the visit covariate and no dropout;
`"A1"` adds informative censoring, `"A7"` is a heavy-switching oncology-style
co-occurrence scenario.

## Estimating both estimands

```{r}
pp  <- va_ct_switch(dat, estimand = "PP",  seed = 1)
itt <- va_ct_switch(dat, estimand = "ITT", seed = 1)

pp
itt
```

The per-protocol effect is typically larger than the intention-to-treat effect:
removing switching un-dilutes the contrast, since control-arm crossover to the
more effective regimen narrows the observed ITT gap.

```{r}
c(ITT = itt$risk_difference, PP = pp$risk_difference)
```

Inference is influence-function based:

```{r}
pp$ci
pp$diagnostics
```

## Your own data

Supply one row per subject with exact times (`Inf` where an event never
happens) and a subject-by-visit covariate matrix:

```{r, eval = FALSE}
dat <- as_va_switch_data(
  subject_data = my_subjects,  # id, A0, W1, W2, T_time, R_time, C_time
  visit_L      = my_visit_L,   # n x length(visit_times)
  visit_times  = c(0, 1, 3, 6, 12),
  tau          = 12
)
va_ct_switch(dat, estimand = "PP")
```

## Heavy switching and positivity

When switching is frequent and strongly covariate-driven, the cumulative
inverse-probability weight can grow large. Percentile truncation on the
correctly specified mechanism controls it while retaining the covariate
dependence:

```{r}
heavy <- simulate_va_switch_trial(n = 1000, scenario = "A7", seed = 2)
va_ct_switch(heavy, estimand = "PP", weight_trunc = "p95", seed = 2)
```

The per-visit weight maxima and effective sample sizes are in
`fit$diagnostics`, supporting the positivity checks that switching-adjustment
guidance recommends.
