---
title: "Testing protocol and current limitations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing protocol and current limitations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

This article gives a short protocol for testing `vacttmle` on trial data.

## Step 1: Install and run smoke tests

```{r install}
install.packages("remotes")
remotes::install_github("blind-contours/vacttmle")
```

```{r smoke}
library(vacttmle)
source(system.file("examples", "trialist-smoke-test.R", package = "vacttmle"))
```

The smoke test should print:

- toy validation rows with `passed = TRUE`
- a VA-CT-TMLE fit with bounded risks and finite standard error
- diagnostics with `converged = TRUE`
- maximum weights of `2` in the baseline-randomized example

## Step 2: Validate your data

Construct a `va_data` object:

```{r data}
library(data.table)

obs <- as_va_data(
  subject_data = subject_data,
  interval_data = interval_data,
  visit_times = c(0, 1, 3, 6, 12),
  tau = 12
)
```

Then confirm:

```{r checks}
validate_va_data(obs)
obs$subject[, .N, by = .(A0, composite_event)]
obs$person_interval[, .N, by = .(k, event_comp)]
```

## Step 3: Run the conservative estimator ladder

Start with g-computation:

```{r gcomp}
fit_gcomp <- va_ct_gcomp(obs, M = 100, seed = 1)
fit_gcomp
```

Then run the proposed estimator:

```{r tmle}
fit_tmle <- va_ct_tmle(
  obs,
  M = 100,
  B = 10,
  max_iter = 10,
  update_method = "adaptive",
  seed = 2
)

fit_tmle
fit_tmle$diagnostics
```

Then run the visit-node ablation:

```{r visit}
fit_visit <- va_visit_tmle(obs, M = 100, B = 10, max_iter = 10, seed = 3)
fit_visit
```

## Step 4: Compare against standard trial summaries

```{r compare}
obs$subject[, .(risk = mean(composite_event), n = .N), by = A0]

survival::survfit(
  survival::Surv(composite_time, composite_event) ~ A0,
  data = obs$subject
)
```

The Kaplan-Meier or Cox summaries are context checks. They do not generally
equal the visit-aligned TMLE estimand when post-baseline covariates and
treatment-history weights matter.

## Step 5: Decide whether the analysis is reportable

A first analysis is ready to share when:

- data validation passes
- endpoint coding matches the scientific question
- all estimates are finite and risks are in `[0, 1]`
- component-wise EIF diagnostics are small or clearly explained
- cumulative weights and ESS are acceptable
- ablation estimates do not reveal an obvious implementation problem

## Current limitations

Supported:

- scheduled visits
- exact event timing inside intervals
- composite endpoint `T or D`
- binary treatment regimes `A = 0` versus `A = 1`
- baseline randomized treatment with deterministic continuation
- known stochastic visit-level treatment probabilities when supplied
- component-wise EIF and weight diagnostics

Not supported:

- hypothetical no-switch/no-D estimands
- treating `D` as censoring
- arbitrary dynamic regimes
- multi-arm or continuous treatment
- recurrent events
- left truncation or delayed entry
- informative censoring before `tau`
- clustered or stratified trial variance estimators
- full Super Learner nuisance libraries

## What to include in an issue

When opening a GitHub issue, include:

- `packageVersion("vacttmle")`
- `sessionInfo()`
- visit grid and `tau`
- endpoint definition
- event counts by arm
- the exact call to `as_va_data()`
- the exact estimator call
- `fit$result`
- `fit$diagnostics`
- whether the smoke test passed locally
