---
title: "Trialist quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trialist quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

This article is for trial analysts using `vacttmle` on scheduled-visit data
with exact event times between visits. The first public workflow estimates the
12-month composite risk difference

```text
{1 - P(T^d1 > tau, D^d1 > tau)} -
{1 - P(T^d0 > tau, D^d0 > tau)}.
```

The intercurrent event `D` is part of the composite endpoint. It is not
censoring in this package version.

## What question is being answered?

For binary treatment, `vacttmle` estimates:

- composite risk by `tau` if everyone followed `A = 0`
- composite risk by `tau` if everyone followed `A = 1`
- risk difference, `A = 1` minus `A = 0`
- EIF standard error and Wald confidence interval
- component-wise EIF diagnostics and cumulative-weight diagnostics

If your scientific question is hypothetical death risk had rescue/crossover not
occurred, do not use this package version for that estimand. That problem treats
the intercurrent event as censoring/adherence and needs a different estimator.

## Data checklist

Your analysis requires one subject table and one person-interval table. The
person-interval table has one row per subject per scheduled visit interval while
the subject remains composite-event-free at the interval start.

Required subject-level columns:

| Column | Requirement |
|---|---|
| `id` | Unique participant id |
| `W1`, `W2` | Baseline covariates in the current interface |
| `A0` | Baseline treatment coded `0` or `1` |
| `T_time` | Failure/event time; `Inf` if not observed before `tau` |
| `D_time` | Intercurrent-event time; `Inf` if not observed before `tau` |
| `composite_time` | `pmin(T_time, D_time, tau)` |
| `composite_event` | `1` if `T` or `D` occurs by `tau`, otherwise `0` |

Required person-interval columns:

| Column | Requirement |
|---|---|
| `id`, `k` | Participant id and interval index, starting at `0` |
| `t_start`, `t_end`, `ell` | Interval start, stop, and length |
| `W1`, `W2`, `A0`, `A_k` | Baseline covariates and interval treatment |
| `L_k`, `L_next` | Visit covariate at interval start and next visit |
| `time_at_risk` | Observed at-risk time inside the interval |
| `event_T`, `event_D` | Cause-specific interval event indicators |
| `event_comp` | Must equal `event_T OR event_D` |
| `event_time` | Elapsed event time within the interval, or `NA` |
| `Y_end` | `1` if composite-event-free at the next visit |

For a baseline-randomized trial with deterministic continuation, the baseline
probability `g0 = 0.5` is enough. For stochastic post-baseline treatment or
management deviations, include known cumulative treatment-history probability
columns in the interval table: `follow_d0`, `follow_d1`, `g_cum_observed`,
`g_cum_d0`, and `g_cum_d1`.

Before fitting, run:

```{r validate}
library(data.table)

obs <- as_va_data(
  subject_data = subject_data,
  interval_data = interval_data,
  visit_times = c(0, 1, 3, 6, 12),
  tau = 12
)

validate_va_data(obs)
```

## Smoke test

Before using your own data, run the installed smoke test:

```{r smoke}
source(system.file("examples", "trialist-smoke-test.R", package = "vacttmle"))
```

The smoke test runs toy validation checks, fits VA-CT-TMLE on simulated
scheduled-visit data, prints the estimate, and prints diagnostics.

## First analysis

Start with the proposed estimator and modest Monte Carlo integration. Increase
`M` later for a final report.

```{r first-fit}
fit <- va_ct_tmle(
  obs,
  d0 = 0,
  d1 = 1,
  tau = 12,
  g0 = 0.5,
  M = 100,
  B = 10,
  max_iter = 10,
  seed = 2026
)

fit
fit$result
fit$diagnostics
```

Example output from the package smoke-test DGM:

```text
Visit-aligned continuous-time estimator
Endpoint: composite
Estimator: ct_tmle
Risk(d0): 0.2875
Risk(d1): 0.1556
Risk difference: -0.1319
SE: 0.0347
95% CI: [-0.1999, -0.0638]
Converged: yes
```

Important result columns:

| Column | Meaning |
|---|---|
| `risk_d0`, `risk_d1` | Composite risk under sustained control and treatment |
| `risk_difference` | `risk_d1 - risk_d0` |
| `se` | EIF standard error for the risk difference |
| `ci_lower`, `ci_upper` | Wald confidence interval |
| `converged` | Whether all targeted EIF components pass the stopping rule |
| `max_abs_eif_T`, `max_abs_eif_D`, `max_abs_eif_Q` | Component-wise EIF means |
| `max_weight_d0`, `max_weight_d1` | Cumulative treatment-history weight maxima |
| `ess_d0`, `ess_d1` | Effective sample size implied by weights |

## Compare with familiar analyses

Use standard trial summaries as context, not as substitutes for the marginal
composite risk difference.

```{r comparisons}
# Composite event counts by arm
obs$subject[, .N, by = .(A0, composite_event)]

# Unadjusted composite risk by randomized arm
obs$subject[, .(risk = mean(composite_event)), by = A0]

# A simple descriptive Cox model for the composite endpoint
survival::coxph(
  survival::Surv(composite_time, composite_event) ~ A0 + W1 + W2,
  data = obs$subject
)
```

Useful comparison questions:

- Are composite-event counts plausible by treatment arm?
- Are adjusted risks close to unadjusted summaries in a balanced randomized
  trial?
- Do VA-CT-TMLE and VA-CT-GCOMP point estimates differ materially?
- Are cumulative weights stable enough for the sustained-regime contrast?
- Did targeting reduce the component-wise EIF means?

## Ablation estimators

Run these as diagnostics:

```{r ablations}
fit_gcomp <- va_ct_gcomp(obs, M = 100, seed = 2027)
fit_visit <- va_visit_tmle(obs, M = 100, B = 10, max_iter = 10, seed = 2028)

fit_gcomp$result
fit_visit$result
```

`va_ct_gcomp()` isolates targeting. `va_visit_tmle()` isolates the added value
of continuous-time hazard targeting.

## What to report

For a first trial report, include:

- package version from `packageVersion("vacttmle")`
- visit grid and target horizon
- endpoint definition, explicitly stating that `D` is part of the composite
- treatment regimes compared
- composite-event counts by arm
- point estimate, SE, and confidence interval
- `fit$diagnostics`
- whether post-baseline treatment probabilities were deterministic or known
  stochastic probabilities
- any non-convergence or extreme weights
